---
title: "Assignment 2"
author: "Miguel Perez Luna"
date: "9/13/2020"
output: 
  html_document:
    theme: cosmo
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```


```{r load libraries}
library(tidyverse)
library(ggplot2)
library(scales)
```

I'll start by loading the person-level data I created in Assignment 1 using read.csv().

```{r}
person_data <- read.csv("transpo_data.csv")
attach(person_data)
```

These are the variables that I am considering: 

1. Sex (SEX) (categorical)

2. Means of Transportation to Work (JWTR) (categorical)

3. Educational Attainment (SCHL) (categorical)

4. Vehicles (1 ton or less) available (vehicle) (continuous)

5. Income (PINCP) (continuous)

6. Travel Time to Work (JWMNP) (continuous)

7. Gross Rent (GRNTP) (continuous)

8. Age (AGEP) (continuous)

I will start by finding the sample mean, sample standard deviation, 95% confidence interval for the population mean, and the interquartile range for my continuous variables. I will also build a histogram for each one to show its distribution. 

## Number of Vehicles Available 

The summary of the number of vehicles is below. 

```{r}
vehicle_summary <- summary(person_data$vehicle)

vehicle_summary

```

The standard deviation of the number of vehicles is below.

```{r}
sd(person_data$vehicle)
```

The 95% confidence interval for the population mean of the number of vehicles is below. 

```{r}
conf_int <- t.test(person_data$vehicle)

conf_int
```

```{r}
ggplot(person_data, aes(x = vehicle)) +
  geom_histogram()
```

```{r}
pdf('vehicle.pdf')
ggplot(person_data, aes(x = vehicle)) +
  geom_histogram()
dev.off()
```



## Income

The summary of the income is below. 

```{r}
PINCP_summary <- summary(person_data$PINCP)

PINCP_summary

```

The standard deviation of the income is below.

```{r}
sd(person_data$PINCP)
```

The 95% confidence interval for the population mean of the income is below. 

```{r}
conf_int <- t.test(person_data$PINCP)

conf_int
```

Mengyao helped me here, she told me about options(scipen = 999) which switched the x-axis from displaying values in scientific notation to displaying values in standard form.

```{r}
options(scipen = 999)
ggplot(person_data, aes(x = PINCP)) +
  geom_histogram()
```

```{r}
pdf('income.pdf')
ggplot(person_data, aes(x = PINCP)) +
  geom_histogram()
dev.off()
```

## Travel Time to Work 

The summary of the travel time to work is below. 

```{r}
JWMNP_summary <- summary(person_data$JWMNP)

JWMNP_summary

```

The standard deviation of the travel time to work is below.

```{r}
sd(person_data$JWMNP)
```

The 95% confidence interval for the population mean of the travel time to work is below. 

```{r}
conf_int <- t.test(person_data$JWMNP)

conf_int
```

```{r}
ggplot(person_data, aes(x = JWMNP)) +
  geom_histogram()
```

```{r}
pdf('traveltime.pdf')
ggplot(person_data, aes(x = JWMNP)) +
  geom_histogram()
dev.off()
```

## Gross Rent

The summary of the gross rent is below. 

```{r}
GRNTP_summary <- summary(person_data$GRNTP)

GRNTP_summary

```

The standard deviation of the gross rent is below.

```{r}
sd(person_data$GRNTP)
```

The 95% confidence interval for the population mean of the gross rent is below. 

```{r}
conf_int <- t.test(person_data$GRNTP)

conf_int
```

```{r}
ggplot(person_data, aes(x = GRNTP)) +
  geom_histogram()
```

```{r}
pdf('rent.pdf')
ggplot(person_data, aes(x = GRNTP)) +
  geom_histogram()
dev.off()
```

## Age

The summary of the age variable is below. 

```{r}
AGEP_summary <- summary(person_data$AGEP)

AGEP_summary

```

The standard deviation of ages is below.

```{r}
sd(person_data$AGEP)
```

The 95% confidence interval for the population mean of ages is below. 

```{r}
conf_int <- t.test(person_data$AGEP)

conf_int
```

```{r}
ggplot(person_data, aes(x = AGEP)) +
  geom_histogram()
```

```{r}
pdf('age.pdf')
ggplot(person_data, aes(x = AGEP)) +
  geom_histogram()
dev.off()
```

## Sex

It looks like there are two possible values of the Sex variable.

```{r}
unique(person_data$SEX_label)
```

The table below shows the number of observations in each category.

```{r}
table(person_data$SEX_label)
```

I'll convert these values to proportions by dividing by the total number of observations. 

```{r}
table(person_data$SEX_label) / sum(table(person_data$SEX_label))
```

The 95% confidence interval for the proportion of females and males is below.

```{r}
t.test(person_data$SEX_label == "Female")
t.test(person_data$SEX_label == "Male")
```


## Means of Transportation to Work

It looks like there are ten possible values of the means of transportation to work variable.

```{r}
unique(person_data$JWTR_label)
```

The table below shows the number of observations in each category.

```{r}
table(person_data$JWTR_label)
```

I'll convert these values to proportions by dividing by the total number of observations. 

```{r}
table(person_data$JWTR_label) / sum(table(person_data$JWTR_label))
```

The 95% confidence interval for the proportion of each value is below.

```{r}
t.test(person_data$JWTR_label == "Bicycle")
t.test(person_data$JWTR_label == "Bus or trolley bus ")
t.test(person_data$JWTR_label == "Car, truck, or van")
t.test(person_data$JWTR_label == "Motorcycle")
t.test(person_data$JWTR_label == "Other method")
t.test(person_data$JWTR_label == "Railroad")
t.test(person_data$JWTR_label == "Streetcar or trolley car (carro publico in Puerto Rico)")
t.test(person_data$JWTR_label == "Subway or elevated")
t.test(person_data$JWTR_label == "Taxicab")
t.test(person_data$JWTR_label == "Walked")
```

## Educational Attainment

It looks like there are twenty three possible values of the educational attainment variable.

```{r}
unique(person_data$SCHL_label)
```

The table below shows the number of observations in each category.

```{r}
table(person_data$SCHL_label)
```

I'll convert these values to proportions by dividing by the total number of observations. 

```{r}
table(person_data$SCHL_label) / sum(table(person_data$SCHL_label))
```

The 95% confidence interval for the proportion of each value is below.

```{r}
t.test(person_data$SCHL_label == "No schooling completed")
t.test(person_data$SCHL_label == "Kindergarten")
t.test(person_data$SCHL_label == "Grade 1")
t.test(person_data$SCHL_label == "Grade 2")
t.test(person_data$SCHL_label == "Grade 3")
t.test(person_data$SCHL_label == "Grade 4")
t.test(person_data$SCHL_label == "Grade 5")
t.test(person_data$SCHL_label == "Grade 6")
t.test(person_data$SCHL_label == "Grade 7")
t.test(person_data$SCHL_label == "Grade 8")
t.test(person_data$SCHL_label == "Grade 9")
t.test(person_data$SCHL_label == "Grade 10")
t.test(person_data$SCHL_label == "Grade 11")
t.test(person_data$SCHL_label == "12th grade - no diploma")
t.test(person_data$SCHL_label == "Regular high school diploma")
t.test(person_data$SCHL_label == "GED or alternative credential")
t.test(person_data$SCHL_label == "Some college, but less than 1 year")
t.test(person_data$SCHL_label == "1 or more years of college credit, no degree")
t.test(person_data$SCHL_label == "Associate's degree")
t.test(person_data$SCHL_label == "Bachelor's degree")
t.test(person_data$SCHL_label == "Professional degree beyond a bachelor's degree")
t.test(person_data$SCHL_label == "Master's degree")
t.test(person_data$SCHL_label == "Doctorate degree")
```









