---
title: "Assignment 1"
author: "Miguel Perez Luna"
date: "9/8/2020"
output: html_document
---

```{r, message=FALSE}
library(tidycensus)
library(tidyverse)
```

## Variables Used:

I decided to explore transportation trends at the person level in the state of Colorado. I selected eight variables from person_vars_2018 and from hh_vars_2018:

1. Sex (SEX) (categorical)

2. Means of Transportation to Work (JWTR) (categorical)

3. Educational Attainment (SCHL) (categorical)

4. Vehicles (1 ton or less) available (VEH) (categorical)

5. Income (PINCP) (continuous)

6. Travel Time to Work (JWMNP) (continuous)

7. Gross Rent (GRNTP) (continuous)

8. Age (AGEP) (continuous)

The VEH variable had several observations that were missing information, and in place of a value expressed "N/A (GQ/vacant)". Furthermore, the same variable expressed numbers of vehicles with labels such as "No vehicles", "1 vehicle", "2 vehicles", and so on up until "6 or more vehicles". I chose to convert this categorical variable to a continuous variable using mutate(). Lastly, I filtered vehicle to only express observations with data, as well as JWMNP to show positive values. I also filtered the other continuous values to only show positive values.

```{r load person-level data, results = 'hide'}
transpo_data <- get_pums(variables = c("SEX",
                                      "JWTR",
                                      "SCHL",
                                      "VEH",
                                      "PINCP",
                                      "JWMNP",
                                      "GRNTP",
                                      "AGEP"),
                        state = "08",
                        year = 2018,
                        survey = "acs1",
                        recode = TRUE) %>%
  mutate(vehicle = case_when(
    VEH_label == "No vehicles" ~ 0,
    VEH_label == "1 vehicle" ~ 1,
    VEH_label == "2 vehicles" ~ 2,
    VEH_label == "3 vehicles" ~ 3,
    VEH_label == "4 vehicles" ~ 4,
    VEH_label == "5 vehicles" ~ 5,
    VEH_label == "6 or more vehicles" ~ 6)) %>%
  filter(vehicle >= 0) %>%
  filter(JWMNP > 0, vehicle > -1, PINCP > -1, GRNTP > -1, AGEP > 0) %>%
  select(SEX_label, JWTR_label, SCHL_label, vehicle, PINCP, JWMNP, GRNTP, AGEP)
```

The first 10 rows of my data are shown below. 

```{r}
transpo_data
```
My final dataset has 6622 rows, which means my analysis will include 6622 persons.

Lastly, I've saved my data as a CSV file.
```{r}
write_csv(transpo_data,"transpo_data.csv")
```