---
title: "Assignment 2"
author: "Miguel Perez Luna"
date: "9/13/2020"
output: 
  html_document:
    theme: cosmo
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

```{r load libraries}
library(tidyverse)
library(ggplot2)
```

I'll start by loading the person-level data I created in Assignment 1 using read.csv().

```{r}
person_data <- read.csv("transpo_data.csv")
attach(person_data)
```

These were the variables that I am considering: 

1. Sex (SEX) (categorical)

2. Means of Transportation to Work (JWTR) (categorical)

3. Educational Attainment (SCHL) (categorical)

4. Vehicles (1 ton or less) available (VEH) (categorical)

5. Income (PINCP) (continuous)

6. Travel Time to Work (JWMNP) (continuous)

7. Gross Rent (GRPIP) (continuous)

8. Age (AGEP) (continuous)

I will start by finding the sample mean, sample standard deviation, 95% confidence interval for the population mean, and the interquartile range for my continuous variables. I will also build a histogram for each one to show its distribution. 

## Income

The summary of the income variable is below. 

```{r}
PINCP_summary <- summary(person_data$PINCP)

PINCP_summary
```

```{r}
paste("The sample mean for income is",
      prettyNum(mean(PINCP), digits = 7), "dollars")

```

